Species richness and network structure jointly drive total biomass and its temporal stability in fish communities

### 2020-08-14 manuscript/bef_stability/manuscript.Rmd
Review discussion

The effect of species richness and network structure on ecosystem properties has
been central to ecology. It has been at the basis of the predictions of the
effects of species richness and network structure on total biomass, productivity
and stability. However, empirical data on the link between network structure and
ecosystem properties are lacking. We used a dataset of standardized fish
sampling in French stream running from 1995 to 2018. We inferred trophic
interactions thanks to recorded body size and diet literature. We then
investigate the direct and indirect effects of stream size and temperature,
species richness and network structure on total biomass and its temporal
stability. We found that species richness and average trophic level had a positive
relationship with total biomass while they had respectively a negative and
positive effect on temporal stability. We also found that species richness and
average trophic level are not independent and that environment had a tremendous
effect on total biomass and its temporal stability notably through its effect on
species richness and average trophic level. Our study highlights that the effects of
species richness and average trophic level should be assessed conjointly and that it
opens exciting perspectives on how environment, species richness and network
structure can jointly determine total biomass in natural communities.

(208/150 words)

@@

#### Biodiversity and ecosystem properties

@@

#### Species richness increases total biomass

@@

#### Multitrophic communities

One has to notice that most of the empirical and experimental evidence of the
positive effect of richness on biomass production and stablity through time
comes from studies focusing on one trophic level, generally primary producers or
primary consumers. Natural ecosystems generally include more than one or two
trophic levels and there is no certainty that the relationship between species
richness and total biomass and its temporal stability for a given trophic level
is representative of this relationship at the food web level. Theoretical
studies showed that food web properties, such as interaction distribution and
strength, affect the distribution of biomass among adjacent trophic levels,
[@thebault_relationship_2006; @schneider_animal_2016; @barbier_pyramids_2019]. A
large body of theoretical studies links the propagation of perturbations to the
foodweb structure [@may_will_1972; @angelis_stability_1975; @dunne_network_2006;
@neutel_reconciling_2007; @duffy_why_2009; @stouffer_compartmentalization_2011],
or trophic chain [@shanafelt_stability_2018; @wang_biodiversity_2016;
@barbier_pyramids_2019]. To date, we have few evidence on the relationship
between network structure and total biomass and its temporal stability in
empirical multitrophic communities.

#### Problematic

@@

#### Environment can mitigate community structure and ecosystem properties

@@

#### Stream fish communities is a good model to study effects of horizontal & vertical diversity

@@

#### Protocol

@@

#### Data selection

@@

#### Biomass

@@

####  Species richness

@@

### Network structure

#### Trophic network inference

```{r}
myload(metaweb_analysis, dir= mypath("data"))
myload(trophic_class, dir= mypath("data", "classes"))
meta <- metaweb_analysis
#metaweb_analysis$resource
#meta$metaweb[meta$resource, meta$resource]
```

Trophic interaction network was inferred from body size and ontogenic diet shift
following previous studies [@gravel_inferring_2013; @brose_predator_2019;
@poisot_structure_2016]. The diet of stream fishes, as many organisms, shifts
over lifespan and thereby over body size. For example, alvins feed on plankton while adults may feed on
algae, fishes, with bigger individuals eating bigger fishes. Each
fish species was hence divided in nine body size classes, each corresponding to a
trophic species (Figure S2), thereby increasing the accuracy of trophic
interactions inference [@jennings_weak_2001]. HERE
Additionally to the fish trophic species, we added seven resource nodes present
in ontogenic food diet database: detritus, biofilm, phytoplankton, zooplankton,
macrophages, phytobenthos and zoobenthos (grey nodes in Figure \@ref(fig:matmet)
C & D). The ontogenic diet database was based on fishbase and literature. Each
fish species had two or three life stages which are delimited by body size
(Figure S2).

The trophic interactions of fish trophic species was determined with the median
of its body size range, hereafter midpoint. The midpoint of trophic species fall
into one of the species life stage and a trophic link was set between the
trophic species and the resources present in the life stage. If the trophic
species is in a piscivorous life stage, the fish-fish trophic links was defined
in two steps and based on the body size ratio between predators and prey
[@mittelbach_ontogeny_1998; @claessen_impact_2002]. First of all, we defined the
range predation window from 3% to 45% of the midpoint of the trophic species.
Then, a trophic link was set between the piscivorous trophic species and every
trophic species whose midpoint was included in the predation range (Figure S2).
Lastly, the trophic links between resources was set according to the food diet
data [@allan_stream_2007; @hart_handbook_2008].

We determined the network structure in two steps: (1) build a metaweb describing the
trophic interactions between all trophic species and the resources (Figure
\@ref(fig:matmet) C), then (2) determine network of a given community by
subtracting the metaweb according to the trophic species present in a given
site and sampling event (Figure \@ref(fig:matmet) D).

For each trophic network, corresponding to a particular site and sampling, we
computed two of the most commonly used network metrics to characterize the food
web, namely the connectance and mean trophic level. Connectance describes 
the level of generalism in the network and is calculated as the number of actual
interaction divided by the maximum number of interaction if all species interact
together ($C = L / N^2$, L and N being resp. the number of
links and of trophic species plus resource nodes). The trophic level of the
nodes was computed of the nodes that the predator nodes consumes
to which is added one, obtained recursively by setting the trophic level of
basal nodes to 1. The mean trophic level $\overline{T}$ was computed as the mean trophic level
weighted by the biomass of each trophic species (excluding resource nodes for
which we had no biomass information), so
$\overline{T} = \sum_i^N{B_iT_i \times \sum_i^N{B_i}^{-1}}$, with $T_i$ and $B_i$
being respectively the trophic level and the biomass of the node $i$.

@@

We characterised the sampled sites by their altitude, slope, distance from
source, stream width and depth and its strahler order, i.e. its position in the
stream network. From the naiades database
[(www.naiades.eaufrance.fr)](http://www.naiades.eaufrance.fr/), we also
got data on water temperature and flow as they affect the type of fish that 
settling down [@franssen_annual_2011; @rolls_scaling_2018; @tonkin_flow_2018;
@gutierrez-canovas_contrasting_2013] as well as BOD (Biological Oxygen Demand)
as a proxy for eutrophisation. As data from the naiades database are not collected
on the same site as our fish sampling sites, we performed an interpolation with
spatial stream networks [@isaak_applications_2014; @hoef_ssn:_2014]. The
interpolation procedure is detailed in supplementary (part 1.2).

We had one environmental variable value for each site and sampling event, except
for water temperature for which the record began in 2006. Then, the temporal
average and CV values of for water temperature, flow, BOD, river width and
depth. As habitat and environmental variables are often collinear, a principal
component analysis (PCA) was performed on the 14 scaled variables. We kept in
the analysis the two first principal components based on the elbow method, on
the eigen values distribution (Fig. S4). The two first principal components
explained `r tot_var_expl_pca`% of the variance. A varimax axis rotation was
performed to maximize the representation of the variables by the principal
components (Figure S4). The two first principal components were used in the
analysis, representing respectively the (1) river size (river width & depth),
(2) altitude, slope and decreasing average temperature. Further description of
the variables are detailed in supplementary information (part 1.3).

@@

#### Summarise SEM statistical framework

Structural Equation Modeling are able to assess multivariate relationships
between variables. It is one promising way to go beyond bivariate relationships
and to investigate complex relationships, direct and
indirect, between variables [@grace_structural_2008; @grace_causal_2014;
@lefcheck_piecewisesem_2016; @grace_integrative_2016]. We built the biomass
model by considering that environment
can directly affect species richness and network structure, as well as total
biomass (Figure \@ref(fig:sem-plot) B). 
Food web structure was shown to be constrained by species richness [@winemiller_must_1989;
@dunne_network_2006; @thebault_stability_2010]. 
We assumed that both species richness and network structure directly affect
total biomass (Figure \@ref(fig:sem-plot) B). In addition,
we assumed that species richness would indirectly affect total biomass through
network structure [@schneider_animal_2016].
We used the same model structure for temporal variability (Figure
\@ref(fig:sem-plot) A). The structural equation models used for total biomass
and its temporal stability contains respectively 4 and 6 response variables, 35
and 26 direct paths.

Structural equation models were based on Gaussian linear models, some variables
were logged to fulfill model assumptions (see residual plots and data
transformation in supplementary informations, Table SXX). To account for the heterogeneity
between the seven hydrographic basins, we set it as random effect on the
intercept of the linear model. The absence of multicollinearity 
was checked with Variance Inflation Factor (VIF, Table SXX). The slopes reported
in the main text were standardized in order to compare their magnitude ($r_\delta
= \beta \frac{\sigma_x}{\sigma_y}$, $\beta$ being the unstandardized slope,
$\sigma_x$ and $\sigma_y$ being respectively the standard deviation of the
predictive variable $x$ and of the response variable $y$). In the main text,
only the slopes which had a $P$ value inferior to $0.05$ were reported (but see
Table SXX and SXX).

@@

Overall, the structural equation models explained a large part of the variation
of connectance, synchrony, $\overline{CV_{sp}}$ and total biomass (resp. $R^2 =
.45$, $.47$, $.49$ and $.38$) but not of mean trophic level ($R^2 =
.18$, Figure \@ref(fig:sem-plot) C and D).

## Total biomass

@@

Species richness and mean trophic level had a significant positive direct effect
on total biomass ($r_\delta = `r rich_on_bm$direct`$ and $r_\delta = 
`r tlvl_on_bm`$ resp., Figure \@ref(fig:sem-plot) D, Table \@ref(tab:indir-sem)),
while connectance had no direct nor indirect significant effect on total biomass (Figure
\@ref(fig:sem-plot) D). The total effect of mean trophic level
on total biomass was of the same magnitude than the one of species richness
(resp. $`r tlvl_on_bm`$ and $`r rich_on_bm$total`$, Table \@ref(tab:indir-sem)).

Species richness had a significant negative effect on mean trophic level
($r_\delta = `r rich_on_tlvl`$) and a strong negative effect on connectance
($r_\delta = `r rich_on_stab$total`$, Figure \@ref(fig:sem-plot) C and D). But,
in turn, connectance had no significant effect on total biomass.

The environmental variables had no significant direct effects on total biomass
(Table SXX), but indirect (Table \@ref(tab:indir-sem)). Average temperature and
altitude respectively increased and decreased total biomass ($r_\delta =
`r env_on_bm$direct["log_RC2"]`$, Table \@ref(tab:indir-sem)), mainly through
species richness ($r_\delta = `r env_on_bm$via_richness["log_RC2"]`$, Table
\@ref(tab:indir-sem)). Average stream size had a positive effect on total
biomass ($r_\delta = `r env_on_bm$total["log_RC1"]`$, Table \@ref(tab:indir-sem)),
equally through species richness and mean trophic level ($r_\delta = 
`r env_on_bm$via_richness["log_RC1"]`$ and $`r env_on_bm$via_com["log_RC1"]`$
resp., Table \@ref(tab:indir-sem)). Average temperature and altitude effects on
total biomass were of the same magnitude than the effects on species richness
and mean trophic level.

## Temporal stability of biomass

By its effect on synchrony and $\overline{CV_{sp}}$, species richness had a
negative total effect on biomass stability ($r_\delta = `r rich_on_stab$total`$,
Table \@ref(tab:indir-sem)). On the one hand, species richness had a direct
negative effect on synchrony ($r_\delta = `r rich_on_sync$direct`$), thereby
increasing stability ($r_\delta = `r rich_on_stab$via_sync`$). On the other
hand, species richness had a
positive direct effect on $\overline{CV_{sp}}$ ($r_\delta = `r rich_on_cv_sp$direct`$),
thereby decreasing biomass stability ($r_\delta = `r rich_on_stab$via_cv_sp`$, Table
\@ref(tab:indir-sem)). Finally, it was surprising that the negative
effect of $\overline{CV_{sp}}$ on temporal stability was twice higher than the
negative effect of synchrony (resp. $r_\delta = `r cv_sp_on_stab`$ and $r_\delta
= `r sync_on_stab`$, Figure \@ref(fig:sem-plot) C).

The mean trophic level had a direct negative effect on
$\overline{CV_{sp}}$ ($r_\delta = `r tlvl_on_cv_sp`$,
Figure \@ref(fig:sem-plot) C) and thereby had a total positive
effect on temporal stability ($r_\delta = `r com_on_stab$via_cv_sp["t_lvl"]`$,
Table \@ref(tab:indir-sem)). The mean trophic level had no significant
effect on synchrony and connectance had no significant direct, nor indirect
effect on temporal stability (Figure \@ref(fig:sem-plot) C, Table SXX).
The total effect of mean trophic level on temporal stability of biomass was of
the same magnitude than the one of species richness (resp.
$`r com_on_stab$total["t_lvl"]`$ and $`r rich_on_stab$total`$, Table
\@ref(tab:indir-sem)).

Average temperature and altitude had a negative and a positive effect on
temporal stability ($r_\delta = `r env_on_stab$total["log_RC2"]`$, Table
\@ref(tab:indir-sem)) mainly through a direct positive effect on
$\overline{CV_{sp}}$
($r_\delta = `r env_on_cv_sp$direct["log_RC2"]`$, Table \@ref(tab:indir-sem)).
Interestingly, average temperature (and opposite of altitude) had also a negative
effect on temporal stability through species richness but a positive one through
mean trophic level 
(resp. $r_\delta = `r env_on_stab$via_richness["log_RC2"]`$ and
$r_\delta = `r env_on_stab$via_com["log_RC2"]`$, Table \@ref(tab:indir-sem)).
This was also true for average river size but effects were weak. Finally, the
effect of average temperature on temporal stability was of higher magnitude than
the effects of species richness and mean trophic level (Table
\@ref(tab:indir-sem)).

```{r indir-stab, cache = FALSE}
caption_sem <- paste0(
  "Direct, indirect and total effect of environment, species richness and
  network structure, i.e. connectance and mean trophic level on total biomass
  and its temporal stability. Indirect effects are computed by multiplying the
  path coefficients from the variable predictor to the response variable, here
  biomass stability. Indirect effects are considered as one step from the target
  variable. For exemple, the indirect effect of species richness on
  biomass stability via its effect on CVsp is $-0.61$. To compute indirect
  effect, only the significant direct effects were taken into account. ",
  "Network structure being the indirect effect that went through both avg trophic
  level and connectance and Species richness_Network structure being the
  indirect effect that went through (i) species richness and from (ii) species
  richness to network structure."
)

@@

Our aim was to evaluate the effects of the network structure on the total
biomass and its temporal stability in natural communities, which have been
poorly documented. To do so, we quantified the effects, directs and indirects,
of habitat properties, species richness, and food web structure, on the total biomass
and stability of fish communities, from the ONEMA fish dataset, across 99
locations in France from 1995 to
2018. We infered trophic interactions networks based on body size-dependency in
food diet from the literature and prey/predator body size ratio for piscivory ^[
We inferred trophic networks at the scale of trophic
species (i.e. nodes being body size classes) to take in account ontogenic diet
shifts and that fish-fish trophic interaction depends on body size ratio between
prey and predator. On the other side, the community metrics were defined at the
species level, (i.e.  species richness, average species CV and synchrony).  ].
We built structural equation models to disantangle the direct and indirect
effects of environmental drivers, species richness and network structure on both
total biomass and biomass stability.  We found that species richness and mean
trophic level had a positive effect on total biomass. In contrast, species
richness had a negative effect on temporal stability whereas mean trophic level
covaried positively.  Futhermore, we found that the effect of species richness
and mean trophic level on total biomass and its temporal stability were not
independant as species richness had a positive effect on mean trophic level.
Finally, we found that the magnitude of the effects of average temperature and
altitude was equal or superior to the effects of species richness and mean
trophic level. Overall, our study is unique by its level of integration, from
environmental effects to total biomass and stability, going through species
richness and network structure, and the first direct assessment of a possible
joint effect of species richness and network structure.

@@

We found that species richness increased total biomass as found in salt marshes,
grasslands and forests [@hector_plant_1999; @tilman_biodiversity_2014;
@wu_relationship_2015; @li_relationship_2018]. The positive effect of species
richness on total biomass might be due to resource use complementarity, which is
held by niche differentiation [@loreau_biodiversity_1998;
@loreau_partitioning_2001] or might be due to a positive selection effect
[@tilman_plant_1997; @loreau_partitioning_2001], which leads to the dominance of
the most productive species. The study of @carey_determining_2011 showed that
resource complementarity might be the main driver in fish communities.

Surprisingly we found a negative relationship between temporal stability and
species richness. Overall, it contrasts with experimental studies in grasslands
[@tilman_biodiversity_2006]; mesocosms (Pennekamp et al. 2018), in 
butterflies natural communities [@olivier_urbanization_2020] or even in lotic
fish communities [@franssen_annual_2011] where positive relationships were found
between temporal stability and species richness. However other studies found a
opposite relationship in natural grassland [@yang_effects_2011] or a marginal
one in natural forests [@declerck_species_2006]. A first step to identify the
mechanisms behind those contrasting patterns is be to look at the effects of
species richness on synchrony and $\overline{CV_{sp}}$, which constitute the two
components of temporal stability quantified as CV [@loreau_species_2008;
@thibaut_understanding_2013].

We found that species richness decreases synchrony, thereby increasing
stability. Species richness can reduce synchrony by decreasing of population
size, thus raising the role of demographic stochasticity on population dynamics
[@loreau_species_2008] but also in the case where species have different
responses to environmental variation [@loreau_species_2008]. While a negative
effect of species richness on synchrony is generally found in experimental
grasslands [@isbell_biodiversity_2009; @roscher_identifying_2011], no relation
has been found in natural forests [@declerck_species_2006], but negative
relationships have been found in natural communities of birds, bats, grasslands,
bats and butterflies [@bluthgen_land_2016; @olivier_urbanization_2020]. Our
study thus highlights that species richness might also decrease synchrony in
empirical trophic communities.

The positive effect of species richness on
$\overline{CV_{sp}}$, thereby decreasing stability, is more intriguing. While
positive effect of species richness on $\overline{CV_{sp}}$ was reported in
experimental grasslands and natural bird communities [@tilman_biodiversity_2006;
@olivier_urbanization_2020], negative and neutral relationships have been found
in butterfly and bat communities. @thibaut_understanding_2013 proposed that a
strongly uneven biomass distribution across population and that the biomass
variance of population increases much faster than its mean can generate a
positive effect of species richness on population ($\overline{CV_{sp}}$)
stability.

>>>>

---
#Our aim was to evaluate the effects of the network structure on the total
#biomass and its temporal stability in natural communities, which have been
#poorly documented. To do so, we quantified the effects, directs and indirects,
#of habitat properties, species richness, and food web structure, on the total biomass
#and stability of fish communities, from the ONEMA fish dataset, across 99
#locations in France from 1995 to
#2018. We infered trophic interactions networks based on body size-dependency in
#food diet from the literature and prey/predator body size ratio for piscivory ^[
#We inferred trophic networks at the scale of trophic
#species (i.e. nodes being body size classes) to take in account ontogenic diet
#shifts and that fish-fish trophic interaction depends on body size ratio between
#prey and predator. On the other side, the community metrics were defined at the
#species level, (i.e.  species richness, average species CV and synchrony).  ].
#We built structural equation models to disantangle the direct and indirect
#effects of environmental drivers, species richness and network structure on both
#total biomass and biomass stability.  
---

Species richness and network structure are at the basis of the predictions of
total biomass, productivity and stability. However, empirical data linking
network structure and ecosystem properties are lacking. We used a dataset of
standardized fish sampling in French streams running from 1995
to 2018. We inferred trophic interactions thanks to recorded body size and diet
literature. We investigated the direct and indirect effects of stream size
and temperature, species richness and network structure on total biomass and its
temporal stability. Species richness and average trophic level had
a positive relationship with total biomass while they had respectively a
negative and positive effect on temporal stability. We found a strong link
between species richness and network structure, highlighting that their effects
should be assessed conjointly. Overall, our study opens fruitful perspectives on
the link between species richness, network structure and ecosystem properties.

(142/150 words)

@@

---
#### Biodiversity and ecosystem properties 
---

@@

---
#### Species richness increases total biomass
---

@@

---
#### Multitrophic communities
---

One has to notice that most of the empirical and experimental evidence of the
positive effect of richness on biomass production and stability through time
comes from studies focusing on one trophic level, generally primary producers or
primary consumers. Natural ecosystems generally include more than one or two
trophic levels and there is no certainty that the relationship between species
richness and total biomass and its temporal stability for a given trophic level
is representative of this relationship at the food web level. Theoretical
studies showed that food web properties, such as interaction distribution and
strength, affect the distribution of biomass among adjacent trophic levels,
[@thebault_relationship_2006; @schneider_animal_2016; @barbier_pyramids_2019]. A
large body of theoretical studies links the propagation of perturbations to the
foodweb structure [@may_will_1972; @angelis_stability_1975; @dunne_network_2006;
@neutel_reconciling_2007; @duffy_why_2009; @stouffer_compartmentalization_2011],
or trophic chain [@shanafelt_stability_2018; @wang_biodiversity_2016;
@barbier_pyramids_2019]. To date, we have few evidence on the relationship
between network structure and total biomass and its temporal stability in
empirical multitrophic communities.

---
#### Problematic
---

@@

---
#### Environment can mitigate community structure and ecosystem properties 
---

@@

---
#### Stream fish communities is a good model to study effects of horizontal & vertical diversity
---

@@

@@

@@

@@

@@

```{r}
myload(metaweb_analysis, dir= mypath("data"))
myload(trophic_class, dir= mypath("data", "classes"))
meta <- metaweb_analysis
#metaweb_analysis$resource
#meta$metaweb[meta$resource, meta$resource]
```
We determined the network structure in two steps: (1) build a metaweb describing the
trophic interactions between all trophic species and the resources (Figure
\@ref(fig:matmet) C), then (2) determine network of a given community by
subtracting the metaweb according to the trophic species present in a given
site and sampling event (Figure \@ref(fig:matmet) D).

Trophic interaction network was inferred from body size and ontogenic diet shift
following previous studies [@gravel_inferring_2013; @brose_predator_2019;
@poisot_structure_2016]. The diet of stream fishes, as many organisms, shifts
over lifespan and thereby over body size. For example, alvins feed on plankton
while adults may feed on algae, fishes, with bigger individuals eating bigger
fishes. Each fish species was hence divided in nine body size classes, each
corresponding to a trophic species (Figure S2). Additionally to the fish trophic
species, we added seven resource nodes present in the ontogenic food diet database:
detritus, biofilm, phytoplankton, zooplankton, macrophages, phytobenthos and
zoobenthos (grey nodes in Figure \@ref(fig:matmet) C & D). The fish trophic
species and the resources constituted the 412 nodes of the metaweb, 45 fish
species divided into nine body classes plus seven resource nodes. The inference
of trophic interactions was shown to be more accurate when considering body size
classes, i.e. trophic species, rather than species [@jennings_weak_2001].

The feeding interactions of a fish node, i.e a trophic species, are determined
by both its species identity and its body size. Then, species identity allows to
differentiate between piscivorous and non piscivorous species, to know on which
resources the trophic species are feeding on and to know the minimum size at
which a piscivorous species begins to feed on fishes. The ontogenic fish diet
database was filled according to fishbase and literature (cf Bonnaffe ref). Each
fish species had two or three life stages which are delimited by a body size
range (Figure S2). The life stage of a fish trophic species was determined with
the center of its body size range, hereafter midpoint (Figure S2).

For a piscivorous trophic species, i.e. a piscivorous node, the fish-fish
trophic links was defined in two steps and based on the body size ratio between
predators and prey. First of all, the predation window of a piscivorous node was
defined as 3% to 45% of its midpoint [@mittelbach_ontogeny_1998;
@claessen_impact_2002; @hart_handbook_2008]. Then, a trophic link was set
between the piscivorous node and every node whose midpoint
was included in the predation range (Figure S2). The fish-resource trophic
links, for piscivorous and non piscivorous nodes, were set between the
trophic species and the resources nodes present in their life stage.
Lastly, the trophic links between resource nodes were set according to the food
diet data [@allan_stream_2007; @hart_handbook_2008].

For each trophic network, corresponding to a particular site and year, we
computed two of the most commonly used network metrics to characterize food
webs, namely the connectance and mean trophic level. Connectance describes 
the level of generalism in the network and is calculated as the number of actual
interaction divided by the maximum number of interaction if all species interact
together ($C = L / N^2$, L and N being resp. the number of
links and of trophic species plus resource nodes). 
The mean trophic level $\overline{T}$ was computed as the mean trophic level
weighted by the biomass of each trophic species (excluding resource nodes for
which we had no biomass information), so $\overline{T} = \sum_i^N{B_iT_i \times
\sum_i^N{B_i}^{-1}}$, with $T_i$ and $B_i$ being respectively the trophic level
and the biomass of the node $i$. The trophic level of the nodes was computed as
one added to the average trophic level of its food item.

@@

We characterised the sites by their altitude, slope, distance from
source, stream width and depth and its strahler order, i.e. its position in the
stream network. From the naiades database
[(www.naiades.eaufrance.fr)](http://www.naiades.eaufrance.fr/), we also
got data on water temperature and flow. As data from the naiades database are
not collected on the same site as our fish sampling sites, we performed an
interpolation with spatial stream networks [@isaak_applications_2014;
@hoef_ssn:_2014]. The interpolation procedure is detailed in supplementary (part
1.2).

We had one environmental variable value for each site and year event, except
for water temperature for which the record began in 2006. Then, we computed the
average and CV values of for water temperature, flow, BOD, river width and
depth. As habitat and environmental variables are often collinear, a principal
component analysis (PCA) was performed on the 14 scaled variables. We kept in
the analysis the two first principal components based on the elbow method, on
the eigen values distribution (Figure S4). The two first principal components
explained `r tot_var_expl_pca`% of the variance (Figure S4). A varimax axis rotation was
performed to maximize the representation of the variables by the principal
components [@kaiser_varimax_1958; @revelle_psych_2019]. The first axis
represented a gradient from small to big streams, characterized by their width
and depth, but also a gradient from stream close to far from the source. The
second axis represented a gradient of increasing average water temperature and
decreasing altitude and slope. Further description of the variables are provided
in supplementary information (part 1.3).

@@

Structural Equation Modeling are able to assess multivariate relationships
between variables. It is one promising way to go beyond bivariate relationships
and to investigate complex relationships, direct and
indirect, between variables [@grace_structural_2008; @grace_causal_2014;
@lefcheck_piecewisesem_2016; @grace_integrative_2016]. The structural equation
model explaining the annual total biomass set that environment
can directly affect species richness and network structure, as well as annual total
biomass (Figure \@ref(fig:sem-plot) B). We assumed that both species richness
and network structure directly affect total biomass (Figure \@ref(fig:sem-plot)
B) and that species richness constrained food web structure
[@winemiller_must_1989; @dunne_network_2006; @thebault_stability_2010].

We used the same model structure for temporal stability of biomass (Figure
\@ref(fig:sem-plot) A). Instead of the annual total biomass, we set the
synchrony and population variability which in turn determined the temporal
stability of biomass. The structural equation models used for annual total
biomass and its temporal stability contains respectively 4 and 6 response
variables, 35 and 26 direct paths.

The structural equation models were based on Gaussian linear models, some variables
were logged to fulfill model assumptions (see residual plots and data
transformation in supplementary informations, Table SXX). We set the seven
hydrographic basins as a random effect on the intercept of the linear models to
account for their heterogeneity. We checked for the absence of multicollinearity 
with Variance Inflation Factor (VIF, Table SXX). The slopes reported
in the main text were standardized in order to compare their magnitude ($r_\delta
= \beta \frac{\sigma_x}{\sigma_y}$, $\beta$ being the unstandardized slope,
$\sigma_x$ and $\sigma_y$ being respectively the standard deviation of the
predictive variable $x$ and of the response variable $y$). The paths whose
slope had a $P$ value inferior to $0.05$ were reported in the main text, all the
path values are presented in Table SXX and SXX).

@@

Overall, the structural equation models explained a large part of the variation
of connectance, synchrony, $\overline{CV_{sp}}$ and total biomass (resp. $R^2 =
.45$, $.47$, $.49$ and $.38$) but less so for mean trophic level ($R^2 =
.18$, Figure \@ref(fig:sem-plot) C and D). We found that species richness had a
significant negative effect on mean trophic level ($r_\delta = 
`r rich_on_tlvl`$) and connectance ($r_\delta = `r rich_on_stab$total`$, Figure
\@ref(fig:sem-plot) C and D), meaning that the effects of species richness and
network structure are not independant.

## Species richness and average trophic level are positively correlated to total biomass

@@

Species richness and mean trophic level had a significant positive direct effect
on total biomass ($r_\delta = `r rich_on_bm$direct`$ and $r_\delta = 
`r tlvl_on_bm`$ resp., Figure \@ref(fig:sem-plot) D, Table \@ref(tab:indir-sem)),
while connectance had no direct significant effect on total biomass (Figure
\@ref(fig:sem-plot) D). The total effect of mean trophic level
on total biomass was of the same magnitude than the one of species richness
(resp. $`r tlvl_on_bm`$ and $`r rich_on_bm$total`$, Table \@ref(tab:indir-sem)).

The two PCA axis capturing environmental variables had no significant direct effects on total biomass
(Table SXX), but indirect (Table \@ref(tab:indir-sem)). The first axis, PCA1,
related to the stream size was positively linked to total biomass ($r_\delta = 
`r env_on_bm$total["log_RC1"]`$, Table \@ref(tab:indir-sem)),
through species richness and mean trophic level too ($r_\delta = 
`r env_on_bm$via_richness["log_RC1"]`$ and $`r env_on_bm$via_com["log_RC1"]`$
resp., Table \@ref(tab:indir-sem)). The second axis, PCA2, related the average
annual temperature and reversely to altitude of the sites was positively related
to total biomass ($r_\delta = `r env_on_bm$total["log_RC2"]`$, Table
\@ref(tab:indir-sem)), mainly through species richness ($r_\delta = 
`r env_on_bm$via_richness["log_RC2"]`$, Table \@ref(tab:indir-sem)). The effect
of PCA2 on total biomass was of the same magnitude than the effects on species
richness and mean trophic level.

## Temporal stability of annual biomass

By its effects on synchrony and $\overline{CV_{sp}}$, species richness had a
negative total effect on biomass stability ($r_\delta = `r rich_on_stab$total`$,
Table \@ref(tab:indir-sem)). On the one hand, species richness had a direct
negative effect on synchrony ($r_\delta = `r rich_on_sync$direct`$), thereby
increasing stability ($r_\delta = `r rich_on_stab$via_sync`$). On the other
hand, species richness had a positive direct effect on $\overline{CV_{sp}}$
($r_\delta = `r rich_on_cv_sp$direct`$),
thereby decreasing biomass stability ($r_\delta = `r rich_on_stab$via_cv_sp`$, Table
\@ref(tab:indir-sem)). The negative effect of $\overline{CV_{sp}}$ on temporal
stability was twice higher than the negative effect of synchrony (resp.
$r_\delta = `r cv_sp_on_stab`$ and $r_\delta = `r sync_on_stab`$, Figure
\@ref(fig:sem-plot) C).

The total effect of mean trophic level on temporal stability of biomass was of
the same magnitude than the one of species richness (resp.
$`r com_on_stab$total["t_lvl"]`$ and $`r rich_on_stab$total`$, Table
\@ref(tab:indir-sem)).
The mean trophic level had only a direct negative effect on
$\overline{CV_{sp}}$ ($r_\delta = `r tlvl_on_cv_sp`$,
Figure \@ref(fig:sem-plot) C) and thereby had a total positive
effect on temporal stability ($r_\delta = `r com_on_stab$via_cv_sp["t_lvl"]`$,
Table \@ref(tab:indir-sem)). Connectance had no significant direct on either
$\overline{CV_{sp}}$ nor synchrony and thus no effect on temporal stability
(Figure \@ref(fig:sem-plot) C, Table SXX).

The total effect of PCA2, related to average annual temperature and reversely to
altitude, on temporal stability was of higher magnitude than the effects of
species richness and mean trophic level (Table
\@ref(tab:indir-sem)).
The PCA2 had a total negative effect on temporal stability 
($r_\delta = `r env_on_stab$total["log_RC2"]`$, Table \@ref(tab:indir-sem)). 
PCA2 had an indirect positive effect on temporal stability through a
direct positive effect on $\overline{CV_{sp}}$ ($r_\delta = 
`r env_on_cv_sp$direct["log_RC2"]`$, Table \@ref(tab:indir-sem)) but also a
negative on through species richness but a positive one through
mean trophic level 
(resp. $r_\delta = `r env_on_stab$via_richness["log_RC2"]`$ and
$r_\delta = `r env_on_stab$via_com["log_RC2"]`$, Table \@ref(tab:indir-sem)).
This was also true for the PCA1, related to stream size, but effects were weaker
than for PCA2.

```{r indir-stab, cache = TRUE, results = TRUE}
caption_sem <- paste0(
  "Direct, indirect and total effect of environment, species richness and
  network structure, i.e. connectance and mean trophic level on total biomass
  and its temporal stability. Indirect effects are computed by multiplying the
  path coefficients from the variable predictor to the response variable, here
  biomass stability. Indirect effects are considered as one step from the target
  variable. For exemple, the indirect effect of species richness on
  biomass stability via its effect on CVsp is $-0.61$. To compute indirect
  effect, only the significant direct effects were taken into account. ",
  "Network structure being the indirect effect that went through both avg trophic
  level and connectance and Species richness_Network structure being the
  indirect effect that went through (i) species richness and from (ii) species
  richness to network structure."
)

@@

We found that species richness and mean
trophic level had a positive effect on total biomass. In contrast, species
richness had a negative effect on temporal stability whereas mean trophic level
covaried positively. Futhermore, we found that the effect of species richness
and mean trophic level on total biomass and its temporal stability were not
independant as species richness had a positive effect on mean trophic level.
Finally, we found that the magnitude of the effects of average temperature and
altitude was equal or superior to the effects of species richness and mean
trophic level. Overall, our study is unique by its level of integration, from
environmental effects to total biomass and stability, going through species
richness and network structure, and the first direct assessment of a possible
joint effect of species richness and network structure.

@@

We found that species richness increased total biomass as found in salt marshes,
grasslands and forests [@hector_plant_1999; @tilman_biodiversity_2014;
@wu_relationship_2015; @li_relationship_2018]. The positive effect of species
richness on total biomass might be due to resource use complementarity [@carey_determining_2011], which is
held by niche differentiation [@loreau_biodiversity_1998;
@loreau_partitioning_2001] or to the higher probability to get high biomass
species  with higher species richness [@loreau_biodiversity_2001].

Surprisingly we found a negative relationship between temporal stability and
species richness. Overall, it contrasts with experimental studies in grasslands
[@tilman_biodiversity_2006]; mesocosms (Pennekamp et al. 2018), in 
butterflies natural communities [@olivier_urbanization_2020] or even in lotic
fish communities [@franssen_annual_2011] where positive relationships were found
between temporal stability and species richness. However other studies found a
opposite relationship in natural grassland [@yang_effects_2011] or a marginal
one in natural forests [@declerck_species_2006].

A first step to identify the mechanisms behind those contrasting patterns is to
look at the effects of species richness on synchrony and $\overline{CV_{sp}}$.
As most studies on the topic, in experimental grasslands
[@isbell_biodiversity_2009; @roscher_identifying_2011], in natural communities
of birds, bats, grasslands, bats and butterflies [@bluthgen_land_2016;
@olivier_urbanization_2020], we found that species richness decreases synchrony,
thereby increasing stability. However, no relation has been found in
natural forests [@declerck_species_2006]. Our study thus
highlights that species richness might also decrease synchrony in empirical
trophic communities. Species richness can reduce synchrony when a given
increase in species richness leads to a decrease of population sizes, which in
turn raises the role of demographic stochasticity on population dynamics
[@loreau_species_2008] but also in the case where species have different
responses to environmental variation [@loreau_species_2008].

The positive effect of species richness on
$\overline{CV_{sp}}$, thereby decreasing stability, is more intriguing. While
positive effect of species richness on $\overline{CV_{sp}}$ was reported in
experimental grasslands and natural bird communities [@tilman_biodiversity_2006;
@olivier_urbanization_2020], negative and neutral relationships have been found
in butterfly and bat communities. However, the ecological mechanisms leading to
such relationships in foodwebs are not known and this question deserves further
research.

---
#@thibaut_understanding_2013 proposed that a
#strongly uneven biomass distribution across populations and that the biomass
#variance of population increases much faster than its mean can generate a
#positive effect of species richness on population ($\overline{CV_{sp}}$)
#stability.
---
